\bf A Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Electrons

### 2025-04-02 TilingProjectDraft/tilingProject.tex
filled in discussions wherever I could, added MFL derivation

\section{Extending the Anderson impurity model}
We have shown in a previous work that by including a local attractive interacting on the conduction bath site coupled to the impurity site of the Anderson impurity model, one can obtain an impurity phase transition from a screened phase to an unscreened local moment phase at a certain critical value of the attractive interaction strength. We now consider a more realistic form of this model, by embedding the impurity site into the lattice of a 2D  conduction bath and accounting for the lowering of the \(s-\)wave symmetry of the interactions down to just the \(C_4\) symmetry of the 2D square lattice. A general structure of such a model is shown in fig.~\ref{embeddedEsiam}; one of the sites of the 2D square lattice is identified as the impurity, while the four nearest-neghbours interact directly with this impurity site. 
\begin{figure}[htpb]
	\includegraphics[width=0.2\textwidth]{pWaveEsiam.pdf}
	\caption{General structure of an embedded extended SIAM, where the impurity site is a part of the 2D square lattice. The four nearest-neghbour sites (dark blue) interact with the impurity site through one or more processes, and can host local interactions. The other lattice sites (light blue) are completely non-interacting.}
	\label{embeddedEsiam}
\subsection{Hamiltonian}
We consider an impurity spin \(\vec S_d\) interacting with a two-dimensional tight-binding conduction bath through a highly localised interaction, described by the Hamiltonian
	\mathcal{H} = H_\text{cbath} + H_\text{imp-cbath} + H_\text{cbath-int}~,
where \(H_\text{cbath}\) is the kinetic energy arising out of nearest-neighbour hopping processes,
	H_\text{cbath} = -2t\sum_{{\bf k}}\left[\cos(ak_x) + \cos(ak_y)\right] c^\dagger_{{\bf k},\sigma}c_{{\bf k},\sigma}~.
The form of the impurity-bath interaction \(H_\text{imp-cbath}\) and the local interaction \(H_\text{cbath-int}\) in the bath can be of various forms, depending in the choice of the bath orbital which participates in the interaction. We now consider some interesting cases:
\paragraph{\(d-\)wave interactions:} This is the case when a coherent \(d-\)wave combination of the bath sites closest to the impurity participates in the interaction:
	d_{\sigma} &= \frac{1}{2}\left(c^\dagger_{L,\sigma} + c^\dagger_{R,\sigma} - c^\dagger_{U\sigma} - c^\dagger_{D\sigma}\right)~,
leading to the interaction terms
	H^{(d)}_\text{imp-cbath} &= J\vec{S}_d\cdot\frac{1}{2}\sum_{\sigma_1,\sigma_2}\vec{\tau}_{\sigma_1,\sigma_2} d^\dagger_{\sigma_1}d_{\sigma_2}~,\\
	H^{(d)}_\text{cbath-int} &= -\frac{W}{2}\left(d^\dagger_{\uparrow}d_{\uparrow} - d^\dagger_{\downarrow}d_{\downarrow}\right)^2~,
where \(\vec \tau\) is the vector of sigma matrices, \(\sigma_1,\sigma_2\) are spin indices, and \(L,R,U,D\) represent the four lattice sites ({\it L}eft, {\it R}ight, {\it U}p, {\it D}own) closest to the impurity. When Fourier-transformed to momentum space, it gives rise to the momentum-dependent couplings
\begin{equation}\begin{aligned}\label{d-wave k-space}
	J_{{\bf k}_1, {\bf k}_2} = J f_{\pi}\left({\bf k}_1\right) f_{\pi}\left({\bf k}_2\right) ~,\quad W_{\left\{ k_i \right\} } = W\prod_{i=1,2,3,4}f_{\pi}(k_i)~,
where \(f_{\pi}({\bf k}) = \cos\left( ak_x \right) - \cos\left( ak_y \right) \) is the \(d-\)wave symmetric functional form in \(k-\)space.
\paragraph{\(p-\)wave interaction:} One can also consider the case when the \(p-\)wave orbital of the nearest sites participates in the interaction:
	p_{\sigma} &= \frac{1}{2}\left(c^\dagger_{L,\sigma} + c^\dagger_{R,\sigma} + c^\dagger_{U\sigma} + c^\dagger_{D\sigma}\right)~,\\
	H^{(p)}_\text{imp-cbath} &= J\vec{S}_d\cdot\frac{1}{2}\sum_{\sigma_1,\sigma_2}\vec{\tau}_{\sigma_1,\sigma_2} p^\dagger_{\sigma_1}p_{\sigma_2}~,\\
	H^{(p)}_\text{cbath-int} &= -\frac{W}{2}\left(p^\dagger_{\uparrow}p_{\uparrow} - p^\dagger_{\downarrow}p_{\downarrow}\right)^2~.
The momentum-dependent Kondo coupling in this case is 
\begin{equation}\begin{aligned}\label{p-wave k-space}
	J_{{\bf k}_1, {\bf k}_2} = J f_{{2\pi}}\left({\bf k}_1\right) f_{{2\pi}}\left({\bf k}_2\right) ~, \quad W_{\left\{ k_i \right\} } = W\prod_{i=1,2,3,4}f_{{2\pi}}(k_i)~,
where \(f_{2\pi}({\bf k}) = \cos\left( ak_x \right) + \cos\left( ak_y \right)\) is the form for the \(p-\)wave symmetric function.

Note that both these interactions involve not only local terms as well as quasi-local terms in the form of off-site interactions that connect the impurity site with multiple bath sites. An example of such a term is \(\vec{S}_d\cdot\frac{1}{2}\sum_{\sigma_1,\sigma_2}\vec{\tau}_{\sigma_1,\sigma_2} c^\dagger_{R,\sigma_1}c_{U,\sigma_2}\), which couples the impurity site with both the right and up sites in the bath. One can consider simple modifications of these cases by retaining only the on-site terms in the interaction Hamiltonian. This gives us the modified \(d-\)wave and \(p-\)wave symmetries:
\begin{equation}\begin{aligned}
	H^{(\bar d(\bar p))}_\text{imp-cbath} &= J\vec{S}_d\cdot\frac{1}{2}\sum_{\sigma_1,\sigma_2}\vec{\tau}_{\sigma_1,\sigma_2} \sum_{i \in \left\{L,R,U,D\right\}} \lambda_i c^\dagger_{i,\sigma_1}c_{i,\sigma_2}~,\\
	H^{(\bar d(\bar p))}_\text{cbath-int} &= -\frac{W}{2}\sum_{i \in \left\{L,R,U,D\right\}} \lambda_i \left(c^\dagger_{i, \uparrow}c_{i, \uparrow} - c^\dagger_{i, \downarrow}c_{i, \downarrow}\right) ~,
\end{aligned}\end{equation}
where \(\lambda_R = \lambda_L = \lambda_U = \lambda_D = 1\) for the \(\bar p\)-wave case, but \(\lambda_R = \lambda_L = -\lambda_U = -\lambda_D = 1\) for the \(\bar d\)-wave case. The \(k-\)dependent couplings have the form
\begin{equation}\begin{aligned}
	J_{{\bf k}_1, {\bf k}_2} &= J f_{\bar d(\bar p)}\left({\bf k}_1 -{\bf k}_2\right),\\
	W_{\left\{ \bf k_i \right\} } &= Wf_{\bar d(\bar p)}\left({\bf k}_1 -{\bf k}_2 + {\bf k}_3 -{\bf k}_4\right) ,
\end{aligned}\end{equation}
where 
\begin{equation}\begin{aligned}
	&f_{\bar d(\bar p)}\left({\bf k}_1 -{\bf k}_2 + {\bf k}_3 -{\bf k}_4\right) \\
	&= \frac{1}{2}\left[\lambda_R \cos\left(k_{1x} - k_{2x} + k_{3x} - k_{4x}\right) \right.\\
	&\quad\left. + \lambda_U \cos\left(k_{1y} - k_{2y} + k_{3y} - k_{4y}\right)\right] ~.
\end{aligned}\end{equation}

\subsection{Unitary RG analysis of the model}
We studied the low-energy physics of this model through a unitary renormalisation group calculation that decouples the high-energy modes of the conduction bath and incorporates their effects in the form of renormalised Hamiltonian couplings. We find that the bath interaction \(W\) remains marginal (does not undergo any renormalisation), while the renormalisation in the momentum-resolved Kondo coupling \(J^{(j)}\) at the \(j^\text{th}\) step is of the form
\begin{equation}\begin{aligned}\label{KondoRGequation}
	\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2} = -\sum_{{\bf q} \in \text{PS}} \frac{J^{(j)}_{{\bf k}_2,{\bf q}} J^{(j)}_{{\bf q},{\bf k}_1} + 4J^{(j)}_{{\bf q}, {\bf \bar q}} W_{{\bf \bar q}, {\bf k}_2, {\bf k}_1, {\bf q}}}{\omega - \frac{1}{2}|\varepsilon_j| + J^{(j)}_{{\bf q}}/4 + W_{{\bf q}}/2}~,
\end{aligned}\end{equation}
where \(\varepsilon_j\) is the energy of the shell being decoupled at the \(j^\text{th}\) step, and the sum is over all momentum states \({\bf q}\) in the particle sector (PS) of the shell \(\varepsilon_j\) (that is, all states that are occupied at \(T=0\) and in the absence of any quantum fluctuations). The Kondo coupling and the RG equation as well have certain symmetries under transformations in the Brillouin zone. If any one of the momenta \({\bf k}_1\) or \({\bf k}_2\) are translated by \(\pi\) (\(k_{1x} \to k_{1x} + \pi, k_{1y} \to k_{1y}+\pi\)), the coupling as well as the RG equation changes sign. Translating both the momenta leads to the reversal of the sign change. 
\begin{equation}\begin{aligned}\label{hamiltonanSymms}
	\Delta J^{(j)}_{{\bf k}_1 + \boldsymbol{\pi}, {\bf k}_2} = \Delta J^{(j)}_{{\bf k}_1, {\bf k}_2 + \boldsymbol{\pi}} = -\Delta J^{(j)}_{{\bf k}_1 + \boldsymbol{\pi}, {\bf k}_2 + \boldsymbol{\pi}} = -\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2}~.
\end{aligned}\end{equation}
These transformations involve translating one or both momenta through the center of the Brillouin zone.

We now consider some specific cases of these results.
\paragraph{Interactions with off-site terms:}
The case with the complete \(d-\)wave and \(p-\)wave terms enjoy a factorisation of the momentum space couplings into functions of the participating momenta, as shown in eq.~\ref{d-wave k-space} and \ref{p-wave k-space}. This factorisation also translates into the RG equation, ensuring that the momentum space Kondo coupling retains its form, and renormalisation occurs only in the magnitude \(J\), through the equation
\begin{equation}\begin{aligned}
	\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2} &= f_{{2\pi}(\pi)}\left({\bf k}_1\right) f_{{2\pi}(\pi)}\left({\bf k}_2\right) \Delta J~,\\
\Delta J &= \sum_{{\bf q} \in \text{PS}} \frac{-f_{2\pi(\pi)}({\bf q})^2\left[J^2 + 4JWf_{2\pi(\pi)}({\bf q})^2\right]}{\omega - \frac{1}{2}|\varepsilon_j| + \frac{J}{4}f_{2\pi(\pi)}({\bf q})^2 + \frac{W}{2}f_{2\pi(\pi)}({\bf q})^4}~.
\end{aligned}\end{equation}
\paragraph{Interactions without off-site terms:}
This choice does not lead to a factorisation of the momentum-space dependence, and the form of the Kondo coupling therefore morphs during the RG flow, allowing the emergence of new degrees of freedom that screen the impurity at low-energies. At the first step, the RG equation can be written down explicitly as
\begin{equation}\begin{aligned}
	&\Delta J^{(0)}_{{\bf k}_1, {\bf k}_2} =\\
	&-\sum_{{\bf q} \in \text{PS}} \frac{J^2 f_{\bar p(\bar d)}\left({\bf k}_2 - {\bf q}\right) f_{\bar p(\bar d)}\left({\bf q} - {\bf k}_1\right) - 4JW f_{\bar p(\bar d)}\left({\bf k}_2 - {\bf k}_1\right)}{\omega - \frac{1}{2}|\varepsilon_j| + \frac{J}{4}f_\mathcal{D}({\bf q})^2 + \frac{W}{4}}
\end{aligned}\end{equation}
The first term in the numerator does not depend only on the momentum difference \({\bf k}_2 - {\bf k}_1\), unlike the bare Kondo coupling. This leads to a new \(k-\)space dependence of the Kondo coupling at the second step, hence leading to a modified RG equation if one makes the \(k-\)dependence explicit.

>>>>

\section{Extending the Anderson Model To a Lattice-Embedded Impurity Model}
Crucial to our method is the choice of an impurity auxiliary model that is able to capture important aspects of the lattice model in a local sense. Some of us have shown in a previous work~\cite{Mukherjee_2023} that almost all aspects of the Mott metal-insulator transition on the infinite dimensional Hubbard model on the Bethe lattice (as seen, for example, from dynamical mean-field theory~\cite{georges1996}) can be captured from a simple extension of the Anderson impurity model (eSIAM). The extension involves including a local attractive interacting \(U_b\) on the conduction bath site coupled to the impurity site; by tuning the ratio of \(|U_b|\) and the Kondo coupling \(J\), we obtained an impurity phase transition from a screened phase to an unscreened local moment phase. Since the impurity site of the auxiliary model is a representative of each correlated site on the lattice, the mapping to the lattice model then involves identifying the locally gapless impurity phase with the bulk gapless of the lattice model. In other words, the ability of the impurity site to participate in low-energy excitations corresponds to the ability of an electron, within the lattice model, to delocalise from one site into the rest of the lattice. In the same vein, a local moment phase of the impurity model maps to a gapped Mott insulating phase on the lattice model, where there's a charge gap to electron hopping processes.

In the present work, the goal is to tackle interacting fermions on a 2D square lattice, which makes it important for us to incorporate the effects of the lattice within the impurity model. More specifically, while the impurity-bath interactions on the infinite dimensional Bethe lattice are \(s-\)wave symmetric, that is not a realistic form for a 2D lattice. To make a realistic impurity model, we embed the impurity site into the lattice of a 2D conduction bath, and lower the \(s-\)wave symmetry of the interactions down to the \(C_4\) symmetry of the 2D square lattice. A general structure of such a model is shown in fig.~\ref{embeddedEsiam}; one of the sites of the 2D square lattice is identified as the impurity (red), while the four nearest-neghbours (navy blue) interact directly with this impurity site through a \(C_4\)-symmetric interaction.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.2\textwidth]{pWaveEsiam.pdf}
	\caption{General structure of an embedded extended SIAM, where the impurity site is a part of the 2D square lattice. The four nearest-neghbour sites (dark blue) interact with the impurity site through one or more processes, and can host local interactions. The other lattice sites (light blue) are completely non-interacting.}
	\label{embeddedEsiam}
\end{figure}
\subsection{Hamiltonian}
We consider an impurity spin \(\vec S_d\) interacting with a two-dimensional tight-binding conduction bath through a highly localised interaction, described by the Hamiltonian
\begin{equation}\begin{aligned}
	\mathcal{H} = H_\text{cbath} + H_\text{imp-cbath} + H_\text{cbath-int}~,
\end{aligned}\end{equation}
where \(H_\text{cbath}\) is the kinetic energy arising out of nearest-neighbour hopping processes,
\begin{equation}\begin{aligned}
	H_\text{cbath} = -2t\sum_{{\bf k}}\left[\cos(ak_x) + \cos(ak_y)\right] c^\dagger_{{\bf k},\sigma}c_{{\bf k},\sigma}~.
\end{aligned}\end{equation}
For the impurity-bath interaction \(H_\text{imp-cbath}\) and the local interaction \(H_\text{cbath-int}\) in the bath, we take the simplest choices consistent with the 2D square lattice geometry:
\begin{equation}\begin{aligned}
	H_\text{imp-cbath} &=\frac{1}{2} J\sum_{\sigma_1,\sigma_2}\sum_{Z} {\bf S}_d\cdot c^\dagger_{Z\sigma_1}{\boldsymbol \tau}_{\sigma_1,\sigma_2} c_{Z\sigma_2}~,\\
	H_\text{cbath-int} &= -\frac{W}{2}\sum_{Z} \left(n_{Z\uparrow} - n_{Z\downarrow}\right)^2 ~,
\end{aligned}\end{equation}
where \(Z\) sums over all nearest-neighbour sites of the impurity site (navy blue circles in Fig.~\ref{embeddedEsiam}, henceforth referred to as bath zeroth sites). The impurity interacts with the bath zeroth sites in \(H_\text{imp-cbath}\) through four local Kondo terms, of uniform strength. The other term \(H_\text{cbath-int}\) represents local Hubbard interactions on the four zeroth sites, again of uniform strength.

The 2D structure of the impurity model and its distinction from the infinite-dimensional counterpart becomes apparent once we transform the Hamiltonian to momentum space:
\begin{equation}\begin{aligned}
	J_{{\bf k}, {\bf k}^\prime} &= \frac{J}{2}\left[\cos\left({\bf k}_x - {\bf k}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y\right)\right] ~,\\
	W_{{\bf k}, {\bf k}^\prime,{\bf q}, {\bf q}^\prime} &= W\left[\cos\left({\bf k}_x - {\bf k}^\prime_x + {\bf q}_x - {\bf q}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y + {\bf q}_y - {\bf q}^\prime_y\right)\right] ~.
\end{aligned}\end{equation}
where the subscripts \(x\) and \(y\) represent components of the momenta vector. This form shows that the Kondo coupling is no longer independent of the exchange momentum, and encodes information of the lattice geometry in it.

\subsection{The unitary renormalisation group method}
In order to obtain the various low-energy phases of our impurity model, we perform a scaling analysis of the associated Hamiltonian using the recently developed unitary renormalisation group (URG) method \cite{anirbanurg1,anirbanurg2}. The method has been applied successfully on a wide variety of problems of correlated fermions~\cite{santanukagome,1dhubjhep,anirbanmott2,anirbanmott1,siddharthacpi,anirban_kondo,Patra_2023,anirbanurg1,anirbanurg2}. The method proceeds by resolving quantum fluctuations in high-energy degrees of freedom, leading to a low-energy Hamiltonian with renormalised couplings and new emergent degrees of freedom. Typically, for a system with Fermi energy \(\epsilon_F\) and bandwidth \(E_N\), the sequence of isoenergetic shells \(\left\{E_{(j)}\right\}, E_{j}\in \left[E_0, E_N\right] \) define the states whose quantum fluctuations we sequentially resolve. The momentum states lying on shells \(E_N\) that are far away from the Fermi surface comprise the UV states, while those on shells near the Fermi surface comprise the IR states. This scheme is shown in Fig.~\eqref{zeromode-bare}.

\begin{figure}[!htb]
	\includegraphics[width=0.3\textwidth]{urgScheme.pdf}
	\caption{High energy - low energy scheme defined and used in the URG method. The states away from the Fermi surface (\(E_N\) in the figure) form the UV subspace and are decoupled first, leading to a Hamiltonian which is more block-diagonal and comprised of only the IR states near the Fermi surface (\(E_0\) in the figure).}
	\label{zeromode-bare}
\end{figure}

As a result of the URG transformations, the Hamiltonian \(H_{(j)}\) at a given RG step \(j\) involves scattering processes between the \(k-\)states that have energies lower than \(D_{(j+1)}\). The unitary transformation \(U_{(j)}\) is then defined so as to remove the number fluctuations of the currently most energetic set of states \(D_{(j)}\)~\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	H_{(j-1)} = U_{(j)} H_{(j)} U^\dagger_{(j)}~, \text{such that} ~\left[H_{(j-1)}, \hat n_{j}\right] =0~.
\end{eqnarray}
The eigenvalue of $\hat{n}_{j}$ has, thus, been rendered an integral of motion (IOM) under the RG transformation.

The unitary transformations can be expressed in terms of a generator \(\eta_{(j)}\) that has fermionic algebra~\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	\label{unitary}
	U_{(j)} = \frac{1}{\sqrt 2}\left(1 + \eta_{(j)} - \eta_{(j)}^\dagger\right)~,~ \quad\left\{ \eta_{(j)},\eta_{(j)}^\dagger \right\} = 1~,
\end{eqnarray}
where \(\left\{\cdot\right\}\) is the anticommutator. The unitary operator \(U_{(j)}\) that appears in Eq.~\eqref{unitary} can be cast into the well-known general form \(U = e^\mathcal{S}, \mathcal{S} = \frac{\pi}{4}\left( \eta^\dagger_{(j)} - \eta_{(j)} \right)\) that a unitary operator can take, defined by an anti-Hermitian operator \(\mathcal{S}\). The generator \(\eta_{(j)}\) is given by the expression~\cite{anirbanurg1,anirbanurg2}
\begin{eqnarray}
	\eta^\dagger_{(j)} = \frac{1}{\hat \omega_{(j)} - \text{Tr}\left(H_{(j)} \hat n_{j}\right) } c^\dagger_{j} \text{Tr}\left(H_{(j)}c_{j}\right)~.
\end{eqnarray}
The operators \(\eta_{(j)},\eta^\dagger_{(j)}\) behave as the many-particle analogues of the single-particle field operators \(c_j,c^\dagger_j\) - they change the occupation number of the single-particle Fock space \(\ket{n_j}\).  The important operator \(\hat \omega_{(j)}\) originates from the quantum fluctuations that exist in the problem because of the non-commutation of the kinetic energy terms and the interaction terms in the Hamiltonian:
\begin{eqnarray}
	\hat \omega_{(j)} = H_{(j-1)} - H^i_{(j)}~.
	\label{omega}
\end{eqnarray}
\(H^i_{(j)}\) is the part of \(H_{(j)}\) that commutes with \(\hat n_j\) but does {\it not} commute with at least one \(\hat n_l\) for \(l < j\). The RG flow continues up to energy \(D^*\), where a fixed point is reached from the vanishing of the RG function. 
Detailed comparisons of the URG with other methods (e.g., the functional RG, spectrum bifurcation RG etc.) can be found in Refs.~\cite{anirbanmott1,anirbanurg1}.

\subsection{Unitary RG analysis of the embedded eSIAM}
We studied the low-energy physics of this model through a unitary renormalisation group calculation that decouples the high-energy modes of the conduction bath and incorporates their effects in the form of renormalised Hamiltonian couplings. We find that the bath interaction \(W\) remains marginal (does not undergo any renormalisation), while the renormalisation in the momentum-resolved Kondo coupling \(J^{(j)}\) at the \(j^\text{th}\) step is of the form
\begin{equation}\begin{aligned}\label{KondoRGequation}
	\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2} = -\sum_{{\bf q} \in \text{PS}} \frac{J^{(j)}_{{\bf k}_2,{\bf q}} J^{(j)}_{{\bf q},{\bf k}_1} + 4J^{(j)}_{{\bf q}, {\bf \bar q}} W_{{\bf \bar q}, {\bf k}_2, {\bf k}_1, {\bf q}}}{\omega - \frac{1}{2}|\varepsilon_j| + J^{(j)}_{{\bf q}}/4 + W_{{\bf q}}/2}~,
\end{aligned}\end{equation}
where \(\varepsilon_j\) is the energy of the shell being decoupled at the \(j^\text{th}\) step, and the sum is over all momentum states \({\bf q}\) in the particle sector (PS) of the shell \(\varepsilon_j\) (that is, all states that are occupied at \(T=0\) and in the absence of any quantum fluctuations). The Kondo coupling and the RG equation as well have certain symmetries under transformations in the Brillouin zone. If any one of the momenta \({\bf k}_1\) or \({\bf k}_2\) are translated by \(\pi\) (\(k_{1x} \to k_{1x} + \pi, k_{1y} \to k_{1y}+\pi\)), the coupling as well as the RG equation changes sign. Translating both the momenta leads to the reversal of the sign change. 
\begin{equation}\begin{aligned}\label{hamiltonanSymms}
	\Delta J^{(j)}_{{\bf k}_1 + \boldsymbol{\pi}, {\bf k}_2} = \Delta J^{(j)}_{{\bf k}_1, {\bf k}_2 + \boldsymbol{\pi}} = -\Delta J^{(j)}_{{\bf k}_1 + \boldsymbol{\pi}, {\bf k}_2 + \boldsymbol{\pi}} = -\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2}~.
\end{aligned}\end{equation}
These transformations involve translating one or both momenta through the center of the Brillouin zone.

With our chosen momentum-space dependence of the Kondo coupling and bath interaction, the form of the Kondo coupling morphs during the RG flow, allowing the emergence of new degrees of freedom that screen the impurity at low-energies. At the first step, the RG equation can be written down explicitly as
\begin{equation}\begin{aligned}
	&\Delta J^{(0)}_{{\bf k}, {\bf k}^\prime} = -\sum_{{\bf q} \in \text{PS}} \frac{J^2 \frac{1}{2}\sum_a\cos\left({\bf k}^\prime_a - {\bf q}_a\right)\frac{1}{2}\sum_a\cos\left({\bf q}_a - {\bf k}_a\right) - 2JW \sum_a\cos\left({\bf k}^\prime_a - {\bf k}_a\right)}{\omega - \frac{1}{2}|\varepsilon_j| + \frac{J}{4}f_\mathcal{D}({\bf q})^2 + \frac{W}{4}}~,
\end{aligned}\end{equation}
where \(a\) sums over the \(x\) and \(y\) components of the momenta. We see that unlike the bare Kondo coupling \(J_{{\bf k}, {\bf k}^\prime}\) that depends only on the momentum difference \({\bf k} - {\bf k}^\prime\), the renormalisation \(\Delta J^{(0)}\) depends on products of the momenta. This shows the form of the Kondo coupling continuously changes under renormalisation.

\begin{figure}[!htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{scattProb-1.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-3.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-4.pdf}
	\includegraphics[width=0.23\textwidth]{scattProb-5.pdf}
	\caption{Variation of the total scattering probability \(\Gamma({\bf k})\) (defined in the main text) as \(W/J\) is increased from left to right. Each plot shows whether \(\Gamma({\bf k})\) is RG-relevant (pink) or irrelevant (black). RG-relevant values of \(\Gamma({\bf k})\) indicate that the \(k-\)point participates in Kondo screening at low-energies, while an irrelevant value indicates that the \(k-\)point is decoupled from the impurity site. As \(W/J\) is made more negative, the antinodal points are the first state to be removed from the Kondo cloud, while the node is the last to be removed.}
	\label{gamma_kxky}
\end{figure}

\subsection{Spin-correlations across the pseudogap}
In figure ~\ref{spinCorr}, we show the evolution of the spin-spin correlation \(\chi_s(d,\vec k) = \left<{\bf S_d}\cdot{\bf S}_{\vec k}\right>\) in \(k-\)space, as the system is tuned through the pseudogap. At vanishing value of the interaction \(W\), the spin correlations are concentrated around the antinode, because of the \(p-\)wave nature of the Kondo interaction. At the entry into the pseudogap, the correlations become more pronounced near the nodes and become suppressed near the antinodes, signalling that the latter parts will gap out first. The last three figures \((W/J=-1.74, -1.83, -1.84)\) show how \(k-\)points starting from the antinode progressively exit the Kondo cloud, the node being the last to decouple from the impurity.

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.23\textwidth]{SF-1.pdf}
	\includegraphics[width=0.23\textwidth]{SF-2.pdf}
	\includegraphics[width=0.23\textwidth]{SF-3.pdf}
	\includegraphics[width=0.23\textwidth]{SF-4.pdf}
	\caption{\(k-\)space distribution of spin-spin correlation \(\chi_s(d,\vec k)\) (defined in the main text). The last three figures show how \(k-\)points starting from the antinode progressively exit the Kondo cloud, the node being the last \(k-\)point to decouple from the impurity.}
	\label{spinCorr}
\end{figure}

\section{Nature of Gapless Excitations in the Impurity Pseudogap}
\subsection{Emergence of a two-channel Kondo Model}
As the bath interaction \(W\) is tuned through the L-PG phase, the four nodal points in the Brillouin zone are the last to decouple from the impurity. This allows us to write down a simpler Kondo model near the transition, focusing on scattering processes involving the nodal region. Specifically, for the Kondo term \(J_{{\bf k}_1, {\bf k}_2})\), we find that only the following classes of scattering processes survive close to the critical point:
\begin{itemize}
	\item Both momenta in a neighbourhood around \({\bf k}_N\), where \({\bf k}_N\) can be any of the four nodal points,
	\item One momentum from the neighbourhood around \({\bf k}_N\), while the other from the neighbourhood around \({\bf k}_N + \left( \pi, \pi \right) \)~.
\end{itemize}
The second class is tied to the first through a symmetry of the Hamiltonian (eq.~\ref{hamiltonanSymms}). The crucial feature of these processes is that each node only interacts with the other node directly opposite to it across the origin. That is, \(\left(\pi/2, \pi/2\right) \) and \(\left(-\pi/2, -\pi/2\right) \) only interact between themselves, while \(\left(\pi/2, -\pi/2\right) \) and \(\left(-\pi/2, \pi/2\right) \) form their own pair of connected regions.

The decoupling of the \(k-\)space regions \(S^\pm\) is captured by the following simplified low-energy Hamiltonian:
\begin{equation}\begin{aligned}\label{twoChannel}
H_\text{eff} = H(S^+) + H(S^-);~H(S) = \sum_{{\bf q}\in S,\sigma}\varepsilon_{\bf q}c^\dagger_{{\bf q}\sigma}c_{{\bf q}\sigma} + \sum_{{\bf q}_1\in S,{\bf q}_2\in S}\sum_{\alpha,\beta}J^*({\bf q}_1,{\bf q}_2){\bf S}_d\cdot {\bf \sigma}_{\alpha\beta}c^\dagger_{{\bf q}_1\alpha}c_{{\bf q}_2\beta}~,
\end{aligned}\end{equation}
where \(H(S^\pm)\) represents the dynamics of momentum states residing in regions \(S^\pm\). The only source of information transfer between the two regions is the fact that they interact with the same impurity local moment. Eq.\ref{twoChannel} tells us that the low-energy dynamics of the embedded eSIAM close to the Kondo-breakdown transition is governed by an emergent {\bf two-channel Kondo model}, with the conduction bath states in the sets \(S^\pm\) making up the two channels.

\subsection{Marginal Fermi liquid behaviour in the pseudogap}
It is well-known that the two-channel Kondo model hosts non-Fermi liquid excitations at low-energies~\cite{Tsvelick_weigmann_mchannel_1985,emery_kivelson}; that this also happens in the present model is corroborated by our calculations of the quasiparticle residue \(Z_\text{imp}\) for the local gapless excitations. For a single-particle Greens function of an interacting model, the quasiparticle residue \(Z\) is defined as the spectral weight of low-energy resonance within the spectral function:
\begin{equation}\begin{aligned}
	G_k = \frac{Z}{\omega - \varepsilon_k - i \Sigma^{\prime\prime}} + G_\text{incoh}~,\quad 1/Z = 1- \frac{d \Sigma^{\prime}}{d\omega}~,
\end{aligned}\end{equation}
where \(\Sigma^\prime\) and \(\Sigma^{\prime\prime}\) are the real and imaginary parts of the self-energy, and \(G_\text{incoh}\) is contribution to the Greens function arising from incoherent short-lived excitations as well as high-energy features that are decoupled from the low-energy excitations~\cite{varma-physrevlett.63.1996}. Small values of \(Z\) indicate that the excitations cannot be described in terms of a Fermi liquid theory. Indeed, as shown in the right panel of Fig.~\ref{channelDecoupling}, the quasiparticle residue \(Z_\text{imp}\) of the impurity excitations becomes vanishingly small in the pseudogap region, signalling that the excitations are of a non-Fermi liquid nature. Given that the pseudogap phase is close to a critical point in the impurity model phase diagram (separating the local Fermi liquid and local moment phases), it is not surprising that a quantum critical model emerges at that point.

Coleman and Ioffe have shown that by mapping the two-channel Kondo problem to a "compactified" form where the impurity spin hybridises with the spin and isospin degrees of freedom of a single conduction channel, the impurity self-energy can be shown to display marginal Fermi liquid behaviour~\cite{Coleman_tsvelik,Schofield1997}:
\begin{equation}\begin{aligned}
	\Sigma \propto \omega \ln \omega~.
\end{aligned}\end{equation}
As some of us have shown in a previous work~\cite{Patra2023MCK}, this can also be obtained from a renormalisation group calculation of the low-energy excitations of the two-channel overscreened fixed point. For the sake of completeness, we have reproduced the calculation here. By considering minimal hopping into the conduction bath, the low-energy excitations above the 2CK ground state are described by the following Hamiltonian~\cite{Patra2023MCK}:
\begin{eqnarray}
	\label{nfl_terms}
	H_\mathrm{eff} =& \frac{8t^2}{{J}}\left[\left(S^z_{0,1}\right)^2 S^+_{0,2} + \left(S^z_{0,2}\right)^2 S^+_{0,1}\right] \left(S^-_{1,1} + S^-_{1,2}\right) + \mathrm{h.c.};
\end{eqnarray}
where \({\boldsymbol S}_{i,l} = \frac{1}{2}\sum_{\sigma_1,\sigma_2}c^\dagger_{i,\sigma_1,l}{\boldsymbol \sigma}_{\sigma_1,\sigma_2}c_{i,\sigma_2,l}\) are spin operators for the \(i^\text{th}\) site on the \(l^\text{th}\) conduction channel (\(l\) can be 1 or 2 for a two-channel model). The labels \(\sigma_1\) and \(\sigma_2\) are spin indices that can be \(\pm 1\). The site index \(0\) represents the bath sites connected immediately to the impurity site. The operator \(c^\dagger_{i,\sigma,l}\) creates an electron with spin \(\sigma\) on the \(i^\text{th}\) site of the \(l^\text{th}\) conduction channel.
